Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)\\ Technical Appendix Including Proofs and Details

### 2017-09-15 fully-abstract-comp/technical-report.tex
moved optional arguments to lstlisting to the correct position (fixed a local auctex problem, so optional arguments no longer cause an issue)

The following pseudo program demonstrates the need of opaque capabilities. If we assume a system with no opaque capabilities, then the following programs break compiler correctness
\begin{lstlisting}
[basicstyle=\sourcecolor{}\ttfamily] 
p1 ::= if r1 is length 2 then
         call r1 with the following callback in r5:
           {put $\textit{\texttt{diverging}}$ closure in r2;
            return};
         halt
       else diverge

@@

The context with $r1$ as an executable capability pointing to:
\begin{lstlisting}
$\scall{r5}{0}$
$\sjmp{r5}$
\end{lstlisting}
\footnote{At this time $\scall{}{}$ did argument spilling and it only takes one argument because we were considering enter capabilities.}distinguishes the two contexts, but two instructions are not enough to do the same at the target level. We would not have enough instructions to set up a proper return pointer for the compiled return to use.

@@

These design goals drove us to the current definition of our source language and some special design choices.
Particularly, contrary to what one might expect in a language with a native stack, we do not use call/return and push/pop instructions.
The reason is that we are working in a language where the length of code blocks can be observable.
Imagine two trusted blocks of code that looks as follows:
\begin{lstlisting}
[basicstyle=\sourcecolor{}\ttfamily] 
p1:       <assert that r1 contains an executable capability to a block of 
           instr_length (push) + instr_length(jmp)>
          <load return capability to p1-ret1 in r2>
          <load value 0 in r3>
          <jump to r1>
p1-ret1:  <assert that 0 is loaded in the next free stack position>
          <repeat the above with value 1 in r3>
          ...
p1-ret2:  <assert that 1 is loaded in the next free stack position>
          <repeat the above with value 2 in r3>
          ...
p1-ret3:  <assert that 2 is loaded in the next free stack position>
          halt

@@

p1 and p2 each accept an (unprotected) closure from the context whose lenght is restricted.
Together with the checks performed by p1 and p2, there is only one possible implementation of the closure that will not fail, namely
\begin{lstlisting}
[basicstyle=\sourcecolor{}\ttfamily] 
closure: push r3
         jmp r2
\end{lstlisting}

>>>>

The following pseudo program demonstrates the need of opaque capabilities. If we assume a system with no opaque capabilities, then the following programs break compiler correctness
\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
p1 ::= if r1 is length 2 then
         call r1 with the following callback in r5:
           {put $\textit{\texttt{diverging}}$ closure in r2;
            return};
         halt
       else diverge

@@

The context with $r1$ as an executable capability pointing to:
\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
$\scall{r5}{0}$
$\sjmp{r5}$
\end{lstlisting}
\footnote{At this time $\scall{}{}$ did argument spilling and it only takes one argument because we were considering enter capabilities.}distinguishes the two contexts, but two instructions are not enough to do the same at the target level. We would not have enough instructions to set up a proper return pointer for the compiled return to use.

@@

These design goals drove us to the current definition of our source language and some special design choices.
Particularly, contrary to what one might expect in a language with a native stack, we do not use call/return and push/pop instructions.
The reason is that we are working in a language where the length of code blocks can be observable.
Imagine two trusted blocks of code that looks as follows:
\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
p1:       <assert that r1 contains an executable capability to a block of 
           instr_length (push) + instr_length(jmp)>
          <load return capability to p1-ret1 in r2>
          <load value 0 in r3>
          <jump to r1>
p1-ret1:  <assert that 0 is loaded in the next free stack position>
          <repeat the above with value 1 in r3>
          ...
p1-ret2:  <assert that 1 is loaded in the next free stack position>
          <repeat the above with value 2 in r3>
          ...
p1-ret3:  <assert that 2 is loaded in the next free stack position>
          halt

@@

p1 and p2 each accept an (unprotected) closure from the context whose lenght is restricted.
Together with the checks performed by p1 and p2, there is only one possible implementation of the closure that will not fail, namely
\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
closure: push r3
         jmp r2
\end{lstlisting}
